<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Working with the Digital Ocean API}
-->

```{r echo=FALSE}
library(analogsea)
knitr::opts_chunk$set(
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```


Working with the Digital Ocean API
======

## Create a DO account

If you don't already have one, [create a DO account](https://www.digitalocean.com/?refcode=0740f5169634). By using this link, you'll start with $10 in credits (enough for >600 hours of computing on a 1 gb machine), and if you become a digital ocean customer we'll get some DO credits for us to offset our costs for testing. Thanks :)

## Authenticate

The best way to authenticate is to generate a [personal access token](https://cloud.digitalocean.com/settings/tokens/new) and save it in an environment variable called `DO_PAT`.  If you don't do this, you'll be prompted to authenticate in your browser the first time you use analogsea.

Make sure you provide digitial ocean your public key at <https://cloud.digitalocean.com/ssh_keys>. Github has some good advice on creating a new public key if you don't already have one: <https://help.github.com/articles/generating-ssh-keys/>.

## Create a droplet

`droplet_create()` will create a droplet on your account. You can run it as below without any inputs, and it will use sensible defaults:

* The smallest memory size of 512mb
* Ubuntu 14.04 box
* Region sfo1
* Uses your ssh key
* Don't use ipv6
* Don't allow backups
* Don't allow private networking

You can set all of thse options in your `.Rprofile` file like `options(do_size = "8gb")` for a default of 8 GB.

The name given to the droplet is picked at random from a list of 1000 random names. 

You can of course set any of these parameters.

```{r}
droplet_create()
```

## Get a droplet or droplets

Listing droplets can be done in singular or plural fashion. `droplet()` accepts a droplet ID, while `droplets()` list all droplets.

If you don't have any droplets yet, you will get an empty list running `droplets()`, and you of course can't pass in a droplet ID number to `droplet()` if you don't have any droplets yet.

```{r eval=FALSE}
droplets()
#> named list()
```

After creating a droplet and running `droplets()` again, we see a list of our droplet(s)

```{r}
(d <- droplets())
```

Or we can pass in a droplet id to `droplet()`

```{r}
droplet(d[[1]]$id)
```

Get more detailed information on your droplet with `summary()`

```{r}
droplet(d[[1]]$id) %>% summary
```

## Actions on droplets

### Delete

You can delete a droplet with `droplet_delete()`. Be careful, as this completely removes your droplet. Backup your droplet or make an image if you want to use the droplet later.

```{r}
droplet_create() %>%
  droplet_delete()
```

### Actions

List actions on a droplet, newer ones at the top. Here's list actions, then rename and list actions again. 

```{r}
droplets()[[1]] %>% droplet_actions()
droplets()[[1]] %>%
  droplet_rename(name = "droppy") %>%
  droplet_wait() %>%
  droplet_actions()
```

### Snapshot

Making a snapshot of a droplet can be done with `droplet_snapshot()`. This action requires that you turn off the droplet first, then take the snapshot. First, create a droplet

```{r}
d <- droplet_create(size = "2gb")
```

Then power off, and take a snapshot, which gives an action object describing that the snapshot is in progress.

```{r}
d %>%
  droplet_power_off() %>%
  droplet_wait() %>%
  droplet_snapshot(name = "mynewsnap")
```

## Regions

The `regions()` function lists region slug names, full names, available sizes, whether the region is available at all, and features. 

```{r}
regions()
```

## Sizes

The `sizes()` function lists size slug names, associated memory, vcpus, disk size, prices, and regions where the size is available. 

```{r}
sizes()
```

## Keys

We suggest you use SSH keys to interact with Digital Ocean from `analogsea`. There are a variety of functions for working with SSH keys. 

List your keys

```{r}
keys()
```

Get a key by id

```{r}
key(keys()[[1]]$id)
```

You can also create a key, rename a key, and delete a key

```{r eval=FALSE}
k <- key_create("key", readLines("~/.ssh/id_rsa.pub"))
k <- key_rename(k, "new_name")
key_delete(k)
```

## Images

The `images()` function can list both your own private images, and public images. If `public=FALSE` only your private images are listed, while if `public=TRUE`, your private images are listed along with publicly avaialble images. 

```{r}
images(public = TRUE, page = 4, per_page = 5)
```

You can also do various actions on images. First, you can pass in an image ID to the `image()` function to get an image object. 

```{r}
img <- images(per_page = 1)[[1]]
image(img$id)
```

You can rename an image

```{r}
img %>% image_rename(name = "analog")
```

You can transfer an image to another region

```{r eval=FALSE}
image(img$id) %>% image_transfer(region = "sfo1")
```

## Domains

You can use domain names for your droplets on Digital Ocean. `analogsea` has a variety of functions to work with domain names. 

List domain names

```{r}
domains()
```

Get a single domain by domain name

```{r}
domain("sckottdrop.info")
```

Create a new domain name

```{r}
dom <- paste0(sample(words, 1), ".info")
domain_create(name = dom, ip_address = "127.0.0.1")
```

Create a domain record, list records and delete the one just created

```{r}
domain("sckottdrop.info") %>% 
  domain_record_create(type = "TXT", name = "hello", data = "world")
records <- domain("sckottdrop.info") %>% domain_records()
domain_record_delete(records[[length(records)]])
```

List records

```{r}
domain("sckottdrop.info") %>% domain_records()
```

Delete a domain name, returns nothing if delete is successful

```{r}
domain(dom) %>% domain_delete()
```

## Docker and Digital Ocean

In the _Docker on Digital Ocean_ vignette we cover working with Docker on Digital Ocean.
